Multi-level generalizations of collective spin dynamics and squeezing

### 2019-10-03 collective_sun/sun_notes.tex
added a bit of discussion for transition and drive operators added coefficients and driving Hamiltonian for a phase-matched Raman drive

The GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}
provide a convenient operator basis to describe dynamics obeying
SU($n$) symmetry.  An external driving field addressing $n$-level
nuclear spins, however, will generally violate SU($n$) symmetry, and
instead obeys the symmetries of the {\it polarization} or {\it
  transition operators} defined by\cite{kryszewski2006alternative,
  bertlmann2008bloch}
\begin{align}
  T_{LM}
  \equiv \sqrt{\f{2L+1}{2I+1}} \sum_{\mu,\nu}
  \bk{I\mu,LM;I\nu} \op{\nu}{\mu},
  \label{eq:trans_ops}
\end{align}
where $L\in\set{0,1,\cdots,n-1}$ and $M\in\set{-L,-L+1,L}$ index a
total spin and its projection onto a quantization axis;
$I\equiv\p{n-1}/2$ is the maximal angular momentum of an $n$-level
spin; $\mu,\nu\in\set{-I,-I+1,\cdots,I}$ index projections of an
$n$-level nuclear spin onto a quantization axis; and
$\bk{I\mu;I\nu,LM}$ is a Clebsch-Gordan coefficient.  Similarly to the
GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
transition operators in \eqref{eq:trans_ops} provide a complete basis
for the space of operators on the Hilbert space of an $n$-level
system, and satisfy the orthonormality condition
\begin{align}
  \obk{T_{LM}|T_{L'M'}} = \delta_{LL'} \delta_{MM'},
\end{align}
which implies
\begin{align}
  H_{\t{int}} = -\f{u}{2} \sum_{p,q} {\v T^{(p)}}^\dag \c \v T^{(q)}
  = -\f{u}{2} \v\T^\dag \c \v\T,
  &&
  \v\T \equiv \sum_p\v T^{(p)},
  \label{eq:H_int_trans}
\end{align}
where $\v T$ is a vector of all transition operators in
\eqref{eq:trans_ops}.

@@

Unlike the GGM operators in \eqref{eq:GGM_diag} and
\eqref{eq:GGM_off_diag}, the transition operators in
\eqref{eq:trans_ops} are not self-adjoint.  In order to express
Hamiltonians in terms of self-adjoint operators, we define the {\it
  drive operators}
\begin{align}
  D_{L,0} \equiv \sqrt{2}~ T_{L,0},
  &&
  D_{LM} \stackrel{M>0}{\equiv} -\p{T_{L\abs{M}} + T_{L\abs{M}}^\dag},
  &&
  D_{LM} \stackrel{M<0}{\equiv} i\p{T_{L\abs{M}} - T_{L\abs{M}}^\dag},
  \label{eq:drive_ops}
\end{align}
for integer $M$ with $\abs{M}\le L$.  Similarly to the GGM operators
in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive
operators are equal to the Pauli operators in the case of $n=2$, and
satisfy the orthonormality condition
\begin{align}
  \obk{D_{LM}|D_{L'M'}} = 2\delta_{LL'}\delta_{MM'},
\end{align}
where $a,b$ index any drive operator in \eqref{eq:drive_ops}.  In
terms of the drive operators,
\begin{align}
  H_{\t{int}} = -\f{u}{4} \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
  = -\f{u}{4} \v\D \c \v\D,
  &&
  \v\D \equiv \sum_p \v D^{(p)},
  \label{eq:H_int_drive}
\end{align}
where $\v D$ is a vector of all drive operators in
\eqref{eq:drive_ops}.

The drive operators have the convenient feature that $D_{1,0}$ and
$D_{1,1}$ are (up to a gauge transformation) respectively proportional
to the Hamiltonians induced on an $n$-level spin by an external
magnetic field and a classical driving laser; loosely speaking, one
can identify $\p{D_{1,0},D_{1,1}}\sim\p{S_\z,S_\x}$, where $S_\alpha$
is a collective SU(2) spin-$\alpha$
operator\cite{perlin2019shorttime}.  More concretely, we can embed the
Hilbert space of a single $n$-level spin into the
permutationally-symmetric (Dicke) manifold of states for $\p{n-1}$
2-level spins via
\begin{align}
  \ket{\mu}_{\t{single}} \to \ket{I+\mu}_{\t{collective}}
  \propto S_+^{I+\mu} \ket\dn^{\otimes\p{n-1}},
\end{align}
where $I\equiv\p{n-1}/2$ is a total spin,
$\mu\in\set{-I,-I+1,\cdots,I}$ indexes a state of the single $n$-level
spin, and $S_+$ is a collective spin-raising operator for $\p{n-1}$
2-level spins.  This embedding takes
\begin{align}
  \p{D_{1,0},D_{1,1},D_{1,-1}}
  \to \p{S_\z,S_\x,S_\y}
  \times \sqrt{\f{3}{I \p{I+1/2} \p{I+1}}}.
\end{align}

@@

An external driving laser addressing a spin-$1/2$ fermion on a lattice
will induce the Hamiltonian
\begin{align}
  \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
  = \Omega \sum_j
  \p{e^{-i\phi_j} c_{j,\up}^\dag c_{j,\dn}
    + e^{i\phi_j} c_{j,\dn}^\dag c_{j,\up}},
  \label{eq:H_drive_2}
\end{align}
where $\Omega$ is a drive strength, $j$ indexes a lattice site, and
$\phi_j$ is a site-dependent phase.  The multi-level analogue of
\eqref{eq:H_drive_2} is
\begin{align}
  H_{\t{drive}}^{(\phi)}
  = -\Omega \sum_j \p{e^{-i\phi_j} T_{1,1}^{(j)}
    + e^{i\phi_j} {T_{1,1}^{(j)}}^\dag}.
  \label{eq:H_drive}
\end{align}
We can make this drive spatially homogenous via the gauge
transformation $c_{j\mu} \to e^{-i\mu\phi_j} c_{j\mu}$, after which the
drive becomes
\begin{align}
  H_{\t{drive}}
  = -\Omega \sum_j\p{T_{1,1}^{(j)} + {T_{1,1}^{(j)}}^\dag}
  = \Omega \sum_j D_{1,1}
  = \Omega \D_{1,1},
\end{align}
where $\D_{LM}\equiv\sum_jD_{LM}^{(j)}$ is a collective drive
operator.  The SU($n$) symmetry of collective interactions implies
that the interaction Hamiltonian $H_{\t{int}}$ in
\eqref{eq:H_int_drive} is invariant under this gauge transformation,
but the single-particle Hamiltonian in \eqref{eq:H_lat} transforms as
\begin{align}
  H_{\t{lat}}
  \to H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} c_{q\mu}^\dag c_{q\mu},
  \label{eq:H_lat_SOC}
\end{align}
where $\phi$ is the relative phase of the driving laser between
adjacent lattice sites.  In the basis of drive operators, this
Hamiltonian takes the form (see Appendix \ref{sec:lat_drive})
\begin{align}
  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L\p{q,\phi} D_{L,0}^{(q)},
  \label{eq:H_lat_drive}
\end{align}
with effective driving field strengths
\begin{align}
  B_L\p{q,\phi} = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
  &&
  f_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Raman drive}
\label{sec:raman}

Rather than coupling states with a single driving laser directly, as
in Section \ref{sec:drive_SOC}, we now consider coupling through an
off-resonant Raman transition to a highly excited state.
Specifically, we consider a three-laser drive of the form
\begin{align}
  H_{\t{Raman,bare}}^{(\phi)}
  = \sum_j \sum_{m\in\set{-1,0,1}} \p{\Omega_m e^{-i\phi_{jm}}
    T_{1,m}^{(j)} \otimes \op{\up}{\dn}^{(j)} + \t{h.c.}}
  + \f{\Delta}{2} \sum_j \1^{(j)}
  \otimes \p{\op{\up}^{(j)} - \op{\dn}^{(j)}},
  \label{eq:raman_bare}
\end{align}
where $j$ indexes a lattice site, $\Omega_m$ is a driving strength;
$\phi_{jm}$ is a site- and drive-dependent phase; $\Delta$ is a
detuning of the lasers from the energy of an excited state; and $\1$
is the identity operator.  If $\Delta\gg\Omega_m$, then at second
order in perturbation theory we can find that the effective action of
$H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state manifold is
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_{\tau\in\set{0,1,2}} H_{\t{Raman},\tau}^{(\phi)}
\end{align}
with
\begin{align}
  H_{\t{Raman},0}^{(\phi)}
  &\equiv -\sum_j\sum_m \f{\Omega_m^2}{\Delta}
  {T_{1,m}^{(j)}}^\dag T_{1,m}^{(j)}, & \label{eq:raman_0} \\
  H_{\t{Raman},1}^{(\phi)}
  &\equiv -\sum_j\sum_{s\in\set{\pm1}}
  \f{\Omega_0\Omega_s}{\Delta} e^{-i\phi_{js}}
  {T_{1,0}^{(j)}}^\dag T_{1,s}^{(j)} + \t{h.c.},
  \\
  H_{\t{Raman},2}^{(\phi)}
  &\equiv -\f{\Omega_1\Omega_{-1}}{\Delta}
  \sum_j\sum_{s\in\set{\pm1}} e^{-i\p{\phi_{js}-\phi_{j,-s}}}
  {T_{1,-s}^{(j)}}^\dag T_{1,s}^{(j)},
  \label{eq:raman_2}
\end{align}
where by choice of gauge we set $\phi_{j,0}=0$, or equivalently take
$\phi_{jm}\to\phi_{jm}-\phi_{j,0}$.  In terms of the drive operators,
we can expand
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_j \sum_{\substack{L\le2\\\abs{m}\le L}}
  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)},
  \label{eq:raman_drive}
\end{align}
where the coefficients $\Omega_{LM}^{(\phi,j)}$ are provided in
Appendix \ref{sec:raman_coeff}.

@@

Here we take the single-particle spin-orbit coupled Hamiltonian
$H_{\t{lat}}^{(\phi)}$ in \eqref{eq:H_lat_SOC}, and derive its
expression in \eqref{eq:H_lat_drive} in terms of the drive operators
$D_{LM}$.  In the basis of ordinary spin operators $s_{\mu\mu}$, this
Hamiltonian takes the form
\begin{align}
  H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} s_{\mu\mu}^{(q)}
  \label{eq:H_lat_spin}
\end{align}
Defining the diagonal ``spin-wave'' operators
\begin{align}
  \omega_{\phi,+} \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
  &&
  \omega_{\phi,-} \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
\end{align}
which are respectively even and odd under spin inversion
($\mu\to-\mu$), we can simplify
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} \omega_{\phi,+} + \sin\p{qa} \omega_{\phi,-},
\end{align}
and resolve the identity (super-)operator $\I$ with
$\I \omega_{\phi,\pm} = \omega_{\phi,\pm}$ in the basis of the drive
operators $D_{LM}$, finding (see Appendix \ref{sec:changing_bases})
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} \f12 \sum_{L,M} \obk{\omega_{\phi,+}|D_{LM}} D_{LM}
  + \sin\p{qa} \f12 \sum_{L,M} \obk{\omega_{\phi,-}|D_{LM}} D_{LM}.
\end{align}
We now note that the drive operators $D_{LM}$ are strictly
off-diagonal for $M\ne0$, and furthermore that $D_{L,0}$ with even
(odd) $L$ is even (odd) under spin inversion, which implies
\begin{align}
  \sum_{L,M} \obk{\omega_{\phi,+}|D_{LM}} D_{LM}
  = \sum_{L\,\t{even}} \obk{\omega_{\phi,+}|D_{L,0}} D_{L,0},
\end{align}
and likewise with $\omega_{\phi,-}$ and odd $L$.  In total, we can
write
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} \f12 \sum_{L\,\t{even}}
  \obk{\omega_{\phi,+}|D_{L,0}} D_{L,0}
  + \sin\p{qa} \f12 \sum_{L\,\t{odd}}
  \obk{\omega_{\phi,-}|D_{L,0}} D_{L,0},
\end{align}
or, more compactly,
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \sum_L B_L\p{q,\phi} D_{L,0},
  \label{eq:compact_SOC_drive}
\end{align}
with effective driving field strengths
\begin{align}
  B_L\p{q,\phi}
  \equiv f_L\p{qa} \f12
  \sum_\mu f_L\p{\mu\phi} \obk{s_{\mu\mu}|D_{L,0}}
  = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
\end{align}
where
\begin{align}
  f_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases}.
\end{align}
Substituting the result in \eqref{eq:compact_SOC_drive} into the
single-particle Hamiltonian in \eqref{eq:H_lat_spin} yields
\begin{align}
  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L\p{q,\phi} D_{L,0}^{(q)}.
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Raman drive coefficients}
\label{sec:raman_coeff}

Here we provide the coefficients $\Omega_{LM}^{(\phi,j)}$ of the
effective Raman driving Hamiltonian
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_j \sum_{\substack{L\in\set{0,1,2}\\\abs{m}\le L}}
  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)}
\end{align}
considered in Section \ref{sec:raman}.  These coefficients are
computed by evaluating products of transition operators in
\eqref{eq:raman_0}--\eqref{eq:raman_2}, using both the structure
constants of the transition operator algebra (derived in Appendix
\ref{sec:trans_prod}) and the fact that
$T_{LM}^\dag = \p{-1}^M T_{L,-M}$.  For clarity, if any coefficient
$\Omega_{LM}\p{\phi}$ is independent of $\phi$, we drop the explicit
dependence on $\phi$, i.e.~taking
$\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.
\begin{align}
  \Omega_{0,0} \equiv -\f1{\sqrt{2n}} \sum_m\f{\Omega_m^2}{\Delta},
  &&
  \Omega_{1,0} \equiv \sqrt{\f{3}{2n\p{n^2-1}}}
  \sum_m m\f{\Omega_m^2}{\Delta},
\end{align}
\begin{align}
  \Omega_{2,0} \equiv -\sqrt{\f{\p{n^2-4}}{10n\p{n^2-1}}}
  \sum_m \p{-1}^m \p{1+\delta_{m,0}} \f{\Omega_m^2}{\Delta},
\end{align}
\begin{align}
  \Omega_{L,1}^{(\phi,j)}
  \equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\phi_{js},
  &&
  \Omega_{L,-1}^{(\phi,j)}
  \equiv \sum_{s\in\set{\pm1}} s~ \omega_{Ls} \sin\phi_{js},
\end{align}
\begin{align}
  \omega_{1,s} \equiv \f{\sqrt3}{\sqrt{n\p{n^2-1}}}
  \f{\Omega_0 \Omega_s}{\Delta},
  &&
  \omega_{2,s} \equiv s~ \f{\sqrt{3/5}}{n\p{n^2-1}}
  \sqrt{\f{\p{n+2}!}{\p{n-3}!}}~ \f{\Omega_0\Omega_s}{\Delta},
\end{align}
\begin{align}
  \Omega_{2,2}^{(\phi,j)}
  \equiv \omega_{2,2} \cos\p{\phi_{j+}-\phi_{j-}},
  &&
  \Omega_{2,-2}^{(\phi,j)}
  \equiv \omega_{2,2} \sin\p{\phi_{j+}-\phi_{j-}},
\end{align}
\begin{align}
  \omega_{2,2} \equiv -\f{\sqrt{6/5}}{n\p{n^2-1}}
  \sqrt{\f{\p{n+2}!}{\p{n-3}!}} \f{\Omega_1\Omega_{-1}}{\Delta}.
\end{align}
Here $\Omega_m$, $\phi_{jm}$, and $\Delta$ are respectively the drive
strengths, phases, and detuning that appear in the ``bare'' Raman
drive in \eqref{eq:raman_bare}.

>>>>

The GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}
provide a convenient operator basis to describe dynamics obeying
SU($n$) symmetry.  An external driving field addressing $n$-level
nuclear spins, however, will generally violate SU($n$) symmetry, and
instead obeys the symmetries of the {\it polarization} or {\it
  transition operators} defined by\cite{kryszewski2006alternative,
  bertlmann2008bloch}
\begin{align}
  T_{LM}
  \equiv \sqrt{\f{2L+1}{2I+1}} \sum_{\mu,\nu}
  \bk{I\mu,LM;I\nu} \op{\nu}{\mu},
  \label{eq:trans_ops}
\end{align}
where $L\in\set{0,1,\cdots,n-1}$ and $M\in\set{-L,-L+1,L}$ index a
total spin and its projection onto a quantization axis;
$I\equiv\p{n-1}/2$ is the maximal angular momentum of an $n$-level
spin; $\mu,\nu\in\set{-I,-I+1,\cdots,I}$ index projections of an
$n$-level nuclear spin onto a quantization axis; and
$\bk{I\mu;I\nu,LM}$ is a Clebsch-Gordan coefficient.  The transition
operator $T_{LM}$ is proportional to the transition induced on a
nuclear spin by the absorption of a spin-$L$ particle with spin
projection $M$ onto a quantization axis.  Similarly to the GGM
operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
transition operators in \eqref{eq:trans_ops} provide a complete basis
for the space of operators on the Hilbert space of an $n$-level
system, and satisfy the orthonormality condition
\begin{align}
  \obk{T_{LM}|T_{L'M'}} = \delta_{LL'} \delta_{MM'},
\end{align}
which implies
\begin{align}
  H_{\t{int}} = -\f{u}{2} \sum_{p,q} {\v T^{(p)}}^\dag \c \v T^{(q)}
  = -\f{u}{2} \v\T^\dag \c \v\T,
  &&
  \v\T \equiv \sum_p\v T^{(p)},
  \label{eq:H_int_trans}
\end{align}
where $\v T$ is a vector of all transition operators in
\eqref{eq:trans_ops}.

@@

Unlike the GGM operators in \eqref{eq:GGM_diag} and
\eqref{eq:GGM_off_diag}, the transition operators in
\eqref{eq:trans_ops} are not self-adjoint.  In order to express
Hamiltonians in terms of self-adjoint operators, we define the {\it
  drive operators}
\begin{align}
  D_{L,0} \equiv \sqrt{2}~ T_{L,0},
  &&
  D_{LM} \stackrel{M>0}{\equiv} -\p{T_{L\abs{M}} + T_{L\abs{M}}^\dag},
  &&
  D_{LM} \stackrel{M<0}{\equiv} i\p{T_{L\abs{M}} - T_{L\abs{M}}^\dag},
  \label{eq:drive_ops}
\end{align}
for integer $M$ with $\abs{M}\le L$.  The drive operator $D_{LM}$ is
proportional to the Hamiltonian induced on a nuclear spin by an
external classical field of spin-$L$ particles with spin projection
$M$ onto a quantization axis.  Similarly to the GGM operators in
\eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive operators
are equal to the Pauli operators in the case of $n=2$, and satisfy the
orthonormality condition
\begin{align}
  \obk{D_{LM}|D_{L'M'}} = 2\delta_{LL'}\delta_{MM'},
\end{align}
which implies
\begin{align}
  H_{\t{int}} = -\f{u}{4} \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
  = -\f{u}{4} \v\D \c \v\D,
  &&
  \v\D \equiv \sum_p \v D^{(p)},
  \label{eq:H_int_drive}
\end{align}
where $\v D$ is a vector of all drive operators in
\eqref{eq:drive_ops}.

The drive operators $D_{1,0}$ and $D_{1,\pm1}$ are proportional to the
Hamiltonians induced on an $n$-level spin by an external magnetic
field and a classical spin-polarized driving laser; loosely speaking,
one can identify $\p{D_{1,0},D_{1,1},D_{1,-1}}\sim\p{S_\z,S_\x,S_\y}$,
where $S_\alpha$ is a collective SU(2) spin-$\alpha$
operator\cite{perlin2019shorttime}.  More concretely, we can embed the
Hilbert space of a single $n$-level spin into the
permutationally-symmetric (Dicke) manifold of states for $\p{n-1}$
2-level spins via
\begin{align}
  \ket{\mu}_{\t{single}} \to \ket{I+\mu}_{\t{collective}}
  \propto S_+^{I+\mu} \ket\dn^{\otimes\p{n-1}},
\end{align}
where $I\equiv\p{n-1}/2$ is a total spin,
$\mu\in\set{-I,-I+1,\cdots,I}$ indexes a state of the single $n$-level
spin, and $S_+$ is a collective spin-raising operator for $\p{n-1}$
2-level spins.  This embedding takes
\begin{align}
  \p{D_{1,0},D_{1,1},D_{1,-1}}
  \to \p{S_\z,S_\x,S_\y}
  \times \sqrt{\f{3}{I \p{I+1/2} \p{I+1}}}.
\end{align}

@@

An driving laser addressing spin-$1/2$ fermions on a 1-D lattice will
induce the Hamiltonian
\begin{align}
  \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
  = \Omega \sum_j
  \p{e^{-i\phi_j} c_{j,\up}^\dag c_{j,\dn}
    + e^{i\phi_j} c_{j,\dn}^\dag c_{j,\up}},
  \label{eq:H_drive_2}
\end{align}
where $\Omega$ is a (real) driving amplitude, $j$ indexes a lattice
site, and $\phi_j$ is a site-dependent phase.  The multi-level
analogue of \eqref{eq:H_drive_2} is
\begin{align}
  H_{\t{drive}}^{(\phi)}
  = -\Omega \sum_j \p{e^{-i\phi_j} T_{1,1}^{(j)}
    + e^{i\phi_j} {T_{1,1}^{(j)}}^\dag}
  = \Omega \D_{1,1}^{(\phi)},
  \label{eq:H_drive}
\end{align}
where
\begin{align}
  \D_{LM}^{(\phi)}
  \equiv \sum_j \sp{\cos\p{M\phi_j} D_{LM}^{(j)}
    + \sin\p{M\phi_j} D_{L,-M}^{(j)}},
  \label{eq:drive_rot}
\end{align}
is an inhomogeneously rotated collective drive operator.  We can make
this drive spatially homogenous via the gauge transformation
$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, after which the
drive becomes simply
\begin{align}
  H_{\t{drive}} = \Omega \D_{1,1}.
\end{align}
The SU($n$) symmetry of collective interactions implies that the
interaction Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_drive} is
invariant under this gauge transformation, but the single-particle
Hamiltonian in \eqref{eq:H_lat} transforms as
\begin{align}
  H_{\t{lat}}
  \to H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} c_{q\mu}^\dag c_{q\mu},
  \label{eq:H_lat_SOC}
\end{align}
where $\phi$ is the relative phase of the driving laser between
adjacent lattice sites.  In the basis of drive operators, this
Hamiltonian takes the form (see Appendix \ref{sec:lat_drive})
\begin{align}
  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L\p{q,\phi} D_{L,0}^{(q)},
  \label{eq:H_lat_drive}
\end{align}
with effective driving field strengths
\begin{align}
  B_L\p{q,\phi} = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
  &&
  f_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Raman drive}
\label{sec:drive_raman}

Rather than coupling the states of a multi-level system directly with
a single driving laser, as in Section \ref{sec:drive_SOC}, we now
consider coupling through an off-resonant Raman transition to a highly
excited state.  Specifically, we consider a three-laser drive of the
form
\begin{align}
  H_{\t{Raman,bare}}^{(\phi)}
  = \sum_j \sum_{m\in\set{1,0,-1}} \p{\Omega_m e^{-i\phi_{jm}}
    T_{1,m}^{(j)} \otimes \op{\up}{\dn} + \t{h.c.}}
  + \f{\Delta}{2} \1 \otimes \p{\op{\up} - \op{\dn}},
  \label{eq:raman_bare}
\end{align}
where $j$ indexes a lattice site, $\Omega_m$ is a (real) driving
amplitude; $\phi_{jm}$ is a site- and drive-dependent phase;
$\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a pseudo-spin operator for
the auxiliary degree of freedom off-resonantly addressed by the
lasers; $\Delta$ is a detuning of the lasers from the energy of an
excited state; and $\1$ is the identity operator.  If the detuning is
large, $\Delta\gg\Omega_m$, then at second order in perturbation
theory we can find that the effective action of
$H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state manifold is
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_{\tau\in\set{0,1,2}} H_{\t{Raman},\tau}^{(\phi)}
\end{align}
with
\begin{align}
  H_{\t{Raman},0}^{(\phi)}
  &\equiv -\sum_j\sum_m \f{\Omega_m^2}{\Delta}
  {T_{1,m}^{(j)}}^\dag T_{1,m}^{(j)}, & \label{eq:raman_0} \\
  H_{\t{Raman},1}^{(\phi)}
  &\equiv -\sum_j\sum_{s\in\set{\pm1}}
  \f{\Omega_0\Omega_s}{\Delta} e^{-i\phi_{js}}
  {T_{1,0}^{(j)}}^\dag T_{1,s}^{(j)} + \t{h.c.},
  \\
  H_{\t{Raman},2}^{(\phi)}
  &\equiv -\f{\Omega_+\Omega_-}{\Delta}
  \sum_j\sum_{s\in\set{\pm1}} e^{-is\p{\phi_{j+}-\phi_{j-}}}
  {T_{1,-s}^{(j)}}^\dag T_{1,s}^{(j)},
  \label{eq:raman_2}
\end{align}
where by choice of gauge we set $\phi_{j,0}=0$, or equivalently
re-define $\phi_{jm}-\phi_{j,0}\to\phi_{jm}$, and we identify the
subscripts $\pm\leftrightarrow\pm1$.  In terms of the drive operators,
we can expand
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_j \sum_{\substack{L\le2\\\abs{M}\le L}}
  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)},
  \label{eq:raman_drive}
\end{align}
where the coefficients $\Omega_{LM}^{(\phi,j)}$ are provided in
Appendix \ref{sec:drive_raman_coeff}.  If the driving lasers are
phase-matched with $\phi_{js}\equiv s\phi_j$, then the drive further
simplifies to
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)}
  \label{eq:drive_raman_matched}
\end{align}
with rotated collective drive operators $\D_{LM}^{(\phi)}$ defined in
\eqref{eq:drive_rot}, and uniform coefficients $\Omega_{LM}$ that have
the scaling
\begin{align}
  \Omega_{0,0} &\propto \Omega_+^2 + \Omega_-^2 + \Omega_0^2,
  &
  \Omega_{1,0} &\propto \Omega_+^2 - \Omega_-^2,
  &
  \Omega_{1,1} &\propto \Omega_0 \p{\Omega_+ + \Omega_-}, \\
  \Omega_{2,0} &\propto \Omega_+^2 + \Omega_-^2 - 2\Omega_0^2,
  &
  \Omega_{2,1} &\propto \Omega_0 \p{\Omega_+ - \Omega_-},
  &
  \Omega_{2,2} &\propto \Omega_+ \Omega_-.
\end{align}
Similarly to the case of a direct drive considered in Section
\ref{sec:drive_SOC}, the Raman drive $H_{\t{Raman}}^{(\phi)}$ in
\eqref{eq:drive_raman_matched} can be made homogenous through the
gauge transformation $c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$,
under which
\begin{align}
  H_{\t{Raman}}^{(\phi)} \to H_{\t{Raman}}
  \equiv \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}.
  \label{eq:drive_raman_homo}
\end{align}

@@

Here we take the single-particle spin-orbit coupled Hamiltonian
$H_{\t{lat}}^{(\phi)}$ in \eqref{eq:H_lat_SOC}, and derive its
expression in \eqref{eq:H_lat_drive} in terms of the drive operators
$D_{LM}$.  In the basis of ordinary spin operators $s_{\mu\mu}$, this
Hamiltonian takes the form
\begin{align}
  H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} s_{\mu\mu}^{(q)}
  \label{eq:H_lat_spin}
\end{align}
Defining the diagonal ``spin-wave'' operators
\begin{align}
  w_{\phi,+} \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
  &&
  w_{\phi,-} \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
\end{align}
which are respectively even and odd under spin inversion
($\mu\to-\mu$), we can simplify
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} w_{\phi,+} + \sin\p{qa} w_{\phi,-},
\end{align}
and resolve the identity (super-)operator $\I$ with
$\I w_{\phi,\pm} = w_{\phi,\pm}$ in the basis of the drive
operators $D_{LM}$, finding (see Appendix \ref{sec:changing_bases})
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} \f12 \sum_{L,M} \obk{w_{\phi,+}|D_{LM}} D_{LM}
  + \sin\p{qa} \f12 \sum_{L,M} \obk{w_{\phi,-}|D_{LM}} D_{LM}.
\end{align}
We now note that the drive operators $D_{LM}$ are strictly
off-diagonal for $M\ne0$, and furthermore that $D_{L,0}$ with even
(odd) $L$ is even (odd) under spin inversion, which implies
\begin{align}
  \sum_{L,M} \obk{w_{\phi,+}|D_{LM}} D_{LM}
  = \sum_{L\,\t{even}} \obk{w_{\phi,+}|D_{L,0}} D_{L,0},
\end{align}
and likewise with $w_{\phi,-}$ and odd $L$.  In total, we can
write
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} \f12 \sum_{L\,\t{even}}
  \obk{w_{\phi,+}|D_{L,0}} D_{L,0}
  + \sin\p{qa} \f12 \sum_{L\,\t{odd}}
  \obk{w_{\phi,-}|D_{L,0}} D_{L,0},
\end{align}
or, more compactly,
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \sum_L B_L\p{q,\phi} D_{L,0},
  \label{eq:compact_SOC_drive}
\end{align}
with effective driving field strengths
\begin{align}
  B_L\p{q,\phi}
  \equiv f_L\p{qa} \f12
  \sum_\mu f_L\p{\mu\phi} \obk{s_{\mu\mu}|D_{L,0}}
  = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
\end{align}
where
\begin{align}
  f_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases}.
\end{align}
Substituting the result in \eqref{eq:compact_SOC_drive} into the
single-particle Hamiltonian in \eqref{eq:H_lat_spin} yields
\begin{align}
  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L\p{q,\phi} D_{L,0}^{(q)}.
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Raman drive coefficients}
\label{sec:drive_raman_coeff}
